Web Service Wrapping, Discovery and Consumption --- More Power to the End-User

### 2010-06-15 trunk/papers/icsoc2010/paper_abstract.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

\begin{abstract}
B2B systems integration was revolutionised by the introduction of web services. The way enterprise systems communicated was dramatically transformed, and by adopting these technologies, the relationships with providers and customers were strengthened. The main advantages seen by companies which adopted the technology were an increase of operational efficiencies, and reduction of costs. 
In this scenario, highly qualified software developers are responsible for the integration of services with other systems, by means of the analysis of WSDL service description or human-readable specifications. 
However this model fails when it tries to target the long tail of enterprise software demand, and as a result, the end-users. Discovery and consumption of web services is far from a straight-forward task for an end-user, meaning potential users, willing to create their task-specific applications, have been ruled out. 
This paper presents an approach to facilitate the discovery and consumption of web services by end-users on the Internet, closing the gap between business web services and the latter. The process includes: (a) a way to generate ready-to-use web services wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs. These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards. Furthermore, two prototypes have been developed to serve as a proof of concept.
\keywords{web services, end-users, discovery, consumption, linked data}
\end{abstract}

>>>>

\begin{abstract}
B2B systems integration was revolutionised by the introduction of web services,
transforming the way enterprise systems communicate and strengthening the  relationships with providers and customers. The main advantage seen by companies was an increase of operational efficiencies, and reduction of costs. 
In this scenario, highly qualified software developers are responsible for the integration of services with other systems, through the analysis of formal service description or human-readable specifications. 
However, this model fails when targeting the long tail of enterprise software demand, the end-users. Discovery and consumption of web services is far from a straight-forward task for an end-user. This means that potential users trying to create task-specific applications are left out. 
This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two. The approach includes: (a) a method to generate ready-to-use web service wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs. 
% These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards. 
\keywords{web services, end-users, discovery, consumption, linked data}
\end{abstract}

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_catalogue.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined. The current state of the art provides several solutions and strategies which providers and consumers may take advantage of. However, as explained in previous sections, a large number of them suffer from limitated syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user. In order to ligthen the issue, this paper presents yet another publishing platform, permiting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving``web service wrappers'' for an easy consumption in any web application.

@@

\begin{itemize}
  \item allows functional discovery through web services pre and postconditions;
	\item serves web services wrappers ready-to-consume in any web application;
	\item provices advanced search capabilites, based on the service formal definition and inferences extracted from it;
	\item supports managing its resources via its RESTful API;
	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
	\item offers web services as Linked Data;
	\item follows well-known best practices for publishing data on the web;
	\item supports content negotiation so the different clients may retrieved the information in JSON, RDF/XML, RDF/N3 or a human-readable HTML version;
\end{itemize}

This platform was intented and it is being used as part of the FAST storyboard tool, and it communicates with others components through its RESTful API, using JSON for the input of the requests. This paper will focus on the functionality regarding web services publication and discovery, and the process to publish any third-party web service into the platform. This process requires two steps: (i) generation of the wrapper (see Sec. \ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform. The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been ommitted, but the main idea is shown):

@@

The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model,
% specified in \ref{ssec:ontology}
acting as a mediator between the public RESTful API and the persistence layer.

@@

% \input{paper_ssec_ontology}

>>>>

The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined. The current state of the art provides several solutions and strategies which providers and consumers may take advantage of. However, as explained in previous sections, a large number of them suffer from limited syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user. In order to lighten the issue, this paper presents yet another publishing platform, permitting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving ``web service wrappers'' for an easy consumption in any web application.

@@

\begin{itemize}
  \item allows functional discovery through web services pre and postconditions;
	\item serves web services wrappers ready-to-consume in any web application;
	\item provides advanced search capabilities, based on the service formal definition and inferences extracted from it;
	\item supports managing its resources via its RESTful API;
	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
	\item offers web services as Linked Data;
	\item follows well-known best practices for publishing data on the web;
	\item supports content negotiation so the different clients may retrieved the information in JSON, RDF/XML, RDF/N3 or a human-readable HTML version;
\end{itemize}

This platform was intended and it is being used as part of the storyboard tool of the FAST platform~\cite{hoyer2009fast}, and it communicates with other components through its RESTful API, using JSON for the input of the requests. This paper will focus on the functionality regarding web services publication and discovery, and the process to publish any third-party web service into the platform. This process requires two steps: (i) generation of the wrapper (see Sec. \ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform. The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been omitted, but the main idea is shown):

@@

The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \ref{ssec:ontology},
acting as a mediator between the public RESTful API and the persistence layer.

@@

\input{paper_ssec_ontology}

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_introduction.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

The remainder of this paper is organised as follows. Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption. In Section 4 we present how we tackle with the web services wrapping, available possibilities and limitations. Next, section 3 describes our platform for publishing web services and empowering end-users to discover services. Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section 6.

>>>>

The remainder of this paper is organised as follows. Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption. In Section 4 we present how we tackle with the web services wrapping, available possibilities and limitations. Next, section 3 describes our platform for publishing web services and empowering end-users to discover services. Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Sect.~\ref{sec:conclusions}.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_use_case.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

The intention of this paper was to show the advances made regarding web service discovery and consumption through the two artifacts developed as part of our research. While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective. Therefore, it is aimed at allowing users (mainly business users at the management level, without programming experience) to create their own situational applications by visually combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs. These two components, as described in Sects. 3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture (see Fig.~\ref{fig:fast_architecture}).

>>>>

The intention of this paper was to show the advances made regarding web service discovery and consumption through the two artifacts developed as part of our research. While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform.
FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective. Therefore, it is aimed at allowing users (mainly business users at the management level, without programming experience) to create their own situational applications by visually combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs. These two components, as described in Sects. 3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture (see Fig.~\ref{fig:fast_architecture}).

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_wrapping.tex
some corrections
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@874 d8b0c59b-6544-0410-94d9-83a5658a058e
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

To define the desired input parameters and the type of the expected result of a new service wrapper the FAST service wrapper tool provides a web based form that allows the editing of these entries, cf. Figure \ref{fig:construct_pre_post_conditions}. Note, we allow to edit example values for the input parameters. These example values may be used to test the service wrapper.

@@

Once the service request is constructed and sent to the service provider, it will send back a response. This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON among others. To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request.

@@

Our tool follows an interactive paradigm. Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, out tool is ontology-driven, therefore, the service wrapper designer shall retrieve the domain-specific types from a FAST ontology server together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding FAST ontology. This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

>>>>

To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web based form that allows the editing of these entries, cf. Figure \ref{fig:construct_pre_post_conditions}. Note, we allow to edit example values for the input parameters. These example values may be used to test the service wrapper.

@@

Once the service request is constructed and sent to the service provider, it will send back a response. This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON. To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request.

@@

Our tool follows an interactive paradigm. Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Sect.~\ref{sec:discovery}) together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Sect.~\ref{ssec:ontology}). This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.
